---
title: "Paper figures (Simulation)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup dir}
library(ctwas)
library(data.table)
source("~/causalTWAS/causal-TWAS/analysis/summarize_basic_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/ld.R")

outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418/"
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210418_compare/"
runtag = "ukb-s80.45-adi"
configtag = 1
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"

ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
exprfs <- read.table(exprfn, header = F, stringsAsFactors = F)[,1]
exprvarfs <- sapply(exprfs, prep_exprvar)

n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes

colorsall <- c("#7fc97f", "#beaed4", "#fdc086")
```

## Parameter estimation figure

In our simulations, the SNP PVE is always set to 0.5 in different settings. The gene PVE is shown as in figures. The number of causal SNPs is always set to 2.5e * $10^{-4}$. In the two settings shown below, number of samples is 45k. For other details about our simulation settings and procedures, please see [here](simulation-ctwas-ukbWG-gtex.adipose_s80.45_mixnormal_041821.html). 

Each plot show one parameter: pi.gene, pi.gene/pi.SNP (enrichment), PVE.gene, PVE.snp. Horizontal bar shows mean true values across the 5 simulations with similar setting parameters. The results by ctwas for each simulation is shown by dots.  

```{r paramplotfunc}
require(latex2exp)
plot_single <- function(mtxlist, truecol, estcol, xlabels = NULL, ...){
  truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
  est <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mtxlist[[x]][, estcol])))
  
  col = est[,1]
  est[,1] <- jitter(est[,1])

  plot(est, pch = 21, xaxt = "n", xlab="" ,frame.plot=FALSE, col= "darkgreen", bg = colorsall[col], ...)
  axis(side=1, at=1, labels = xlabels, tick = F)
  #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
  for (t in 1:nrow(truth)){
    row <- truth[t,]
    segments(row[1]-0.2, row[2] , row[1] + 0.2, row[2],
       col = colorsall[t], lty = par("lty"), lwd = 2, xpd = FALSE)
  }
  grid()
}
 
get_params <- function(configtag, runtag, simutaglist){
    mtxlist <- list()
  for (group in 1:length(simutaglist)){
    simutags <- simutaglist[[group]]
    source(paste0(outputdir, "config", configtag, ".R"))
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
    susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
    mtxlist[[group]] <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  }
    return(mtxlist)
}

plot_PIP <- function(configtag, runtag,  simutags, ...){
   phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

   f1 <- caliPIP_plot(phenofs, susieIfs, ...) 
   return(f1)
}
```

```{r paramplot, fig.width= 8, fig.height= 18}
simutaglist = lapply(1:10, function(x) paste(x, 1:5, sep ="-"))
mtxlist <- get_params(configtag, runtag, simutaglist)
pdf(file = "temp.pdf", height = 25, width =5)
par(mfrow=c(length(simutaglist),4))
for (i in 1:length(simutaglist)){
  plot_single(mtxlist[i], truecol = "enrich_truth", estcol = "enrich_est", main = TeX("Enrichment $\\pi_G/\\pi_V$"), ylim = c(0,200), xlim = c(0.8,1.2))
  plot_single(mtxlist[i], truecol = "pi1.gene_truth", estcol = "pi1.gene_est", main = TeX("Percent causal, $\\pi_G$"), ylim = c(0,0.1), xlim = c(0.8,1.2))
   plot_single(mtxlist[i], truecol = "PVE.gene_truth", estcol = "PVE.gene_est", main = "Gene PVE", ylim = c(0,1), xlim = c(0.8,1.2))
    plot_single(mtxlist[i], truecol = "PVE.SNP_truth", estcol = "PVE.SNP_est",main = "SNP PVE", ylim = c(0,1), xlim = c(0.8,1.2))
}
dev.off()
```


## PIP Figure
```{r PIPplot, fig.width= 8, fig.height= 18}
simutaglist = c(lapply(1:4, function(x) paste(x, 1:5, sep ="-")),
                list(paste(5, c(1,3,4,5), sep = "-")),
                lapply(6:10, function(x) paste(x, 1:5, sep ="-")))
pdf(file = "temp2.pdf", height = 25, width =3)
plotlist <- list()
for (i in 1:length(simutaglist)){
  # for (i in 1:2){
  plotlist[[i]] <- plot_PIP(configtag, runtag, simutaglist[[i]], main = "")
}
gridExtra::grid.arrange(plotlist[[1]], plotlist[[2]],
                        plotlist[[3]], plotlist[[4]],
                        plotlist[[5]], plotlist[[6]],
                        plotlist[[7]], plotlist[[8]],
                        plotlist[[9]], plotlist[[10]],
                        ncol =1)
dev.off()
```

## Regions plot





